A seamless, client-centric programming model for type safe web applications

### 2014-05-08 paper.tex
typo
Kill use/RemoteValue

While explaining of the properties of our solution is all well and good,
nothing compares to a good old Hello World example to convey the idea.
We begin by implementing a function which prints a greeting to the server's
console.

@@

\begin{code}
liftServerIO :: IO a -> App (RemoteValue a)
\end{code}

In contrast to \lstinline!liftIO!, the result value of \lstinline!liftServerIO!
is wrapped in the abstract \lstinline!RemoteValue! type. As the name implies,
values of this type live exclusively on the server. In order to inspect such a
value is needs to be extracted somehow, something which is only possible using
a single function:

\begin{code}
use :: RemoteValue a -> Server a
\end{code}

Therein lies the secret to restricting \lstinline!liftServerIO! to the
server: the only function capable of turning a \lstinline!RemoteValue! into
one that can be inspected lives in the server monad. Any client side code is
thus free to completely ignore executing computations lifted using
\lstinline!liftServerIO!; since the result of a server lifted computation is
never observable on the client, the client has no obligation to even produce
such a value. Figure \ref{lst:good-state} shows how to make proper use of
server side state.

@@

  count <- remote $ do
    r <- use remoteref
    liftIO $ atomicModifyIORef r (\v -> (v+1, v+1))

@@

\begin{code}
srvHello :: Remote RcptList -> Server ()
srvHello remoteRcpts = do
  recipients <- use remoteRcpts
  sid <- getSessionID
  liftIO $ do
    rcptMVar <- CC.newEmptyMVar
    CC.modifyMVar recipients $ \cs ->
      return ((sid, rcptMVar):cs ,())
\end{code}

@@

\begin{code}
srvSend :: Remote RcptList -> String -> Server ()
srvSend remoteRcpts message = do
    rcpts <- use remoteRcpts
    liftIO $ do
      recipients <- CC.readMVar rcpts
      mapM_ (CC.forkIO . deliver message) recipients
  where
    deliver msg (_,rcptMVar) = CC.putMVar rcptMVar msg
\end{code}

@@

\begin{code}
srvAwait :: Remote RcptList -> Server String
srvAwait remoteRcpts = do
  rcpts <- use remoteRcpts
  sid <- getSessionID
  liftIO $ do
    recipients <- CC.readMVar rcpts
    case lookup sid recipients of
      Just mv -> CC.takeMVar mv
      _       -> fail "Unregistered session!"
\end{code}

@@

use          :: RemoteValue a -> Server a
getSessionID :: Server SessionID
\end{code}
\caption{Types of the Haste.App core functions}
\label{lst:api}
\end{listingfloat}

@@

\begin{code}
data RemoteValue a = Here a
data Remote a = Remote
\end{code}

Where the client is able to do something useful with the \lstinline!Remote!
type but not with \lstinline!RemoteValue! the server has no way to inspect
\lstinline!Remote! functions, and thus only has a no-op implementation of the
\lstinline!<.>! operator. On the other hand, the server has full access to the
contents of \lstinline!RemoteValue!s via the \lstinline!use! function.

\begin{code}
use :: RemoteValue a -> Server a
use (Here x) = return x
\end{code}

Those values are produced by the \lstinline!remoteIO! function, which is quite
simple. It merely executes its argument and wraps its argument in the
\lstinline!RemoteValue! type.

\begin{code}
remoteIO :: IO a -> Server (RemoteValue a)
remoteIO m = App $ fmap Here m
\end{code}

>>>>

While explaining the properties of our solution is all well and good,
nothing compares to a good old Hello World example to convey the idea.
We begin by implementing a function which prints a greeting to the server's
console.

@@

\begin{code}
liftServerIO :: IO a -> App (Server a)
\end{code}

In contrast to \lstinline!liftIO!, the result value of \lstinline!liftServerIO!
is contained within the \lstinline!Server! monad, completely out of reach of the
client. Any client side code is thus free to completely ignore executing
computations lifted using \lstinline!liftServerIO!; since the result of a
server lifted computation is never observable on the client, the client has no
obligation to even produce such a value. Figure \ref{lst:good-state} shows how
to make proper use of server side state.

@@

  count <- remote $ do
    r <- remoteref
    liftIO $ atomicModifyIORef r (\v -> (v+1, v+1))

@@

\begin{code}
srvHello :: Server RcptList -> Server ()
srvHello remoteRcpts = do
  recipients <- remoteRcpts
  sid <- getSessionID
  liftIO $ do
    rcptMVar <- CC.newEmptyMVar
    CC.modifyMVar recipients $ \cs ->
      return ((sid, rcptMVar):cs ,())
\end{code}

@@

\begin{code}
srvSend :: Server RcptList -> String -> Server ()
srvSend remoteRcpts message = do
    rcpts <- remoteRcpts
    liftIO $ do
      recipients <- CC.readMVar rcpts
      mapM_ (CC.forkIO . deliver message) recipients
  where
    deliver msg (_,rcptMVar) = CC.putMVar rcptMVar msg
\end{code}

@@

\begin{code}
srvAwait :: Server RcptList -> Server String
srvAwait remoteRcpts = do
  rcpts <- remoteRcpts
  sid <- getSessionID
  liftIO $ do
    recipients <- CC.readMVar rcpts
    case lookup sid recipients of
      Just mv -> CC.takeMVar mv
      _       -> fail "Unregistered session!"
\end{code}

@@

getSessionID :: Server SessionID
\end{code}
\caption{Types of the Haste.App core functions}
\label{lst:api}
\end{listingfloat}

@@

\begin{code}
data Remote a = Remote
\end{code}

Where the client is able to do something useful with the \lstinline!Remote!
type, the server has no way to inspect \lstinline!Remote! functions, and thus
only has a no-op implementation of the \lstinline!<.>! operator. On the other
hand, it does have full access to the values and side effects produced by
the \lstinline!liftServerIO! function, which is quite simple. The side effect
is performed immediately and its result saved for later, as state holding
elements should only be initialized once.

\begin{code}
liftServerIO :: IO a -> App (Server a)
liftServerIO m = App $ do
  x <- liftIO m
  return (return x)
\end{code}
